---
title: "Getting and Cleaning Data Course Project"
author: "Andre Morato"
date: "5 de abril de 2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

tBodyAcc-XYZ
tGravityAcc-XYZ
tBodyAccJerk-XYZ
tBodyGyro-XYZ
tBodyGyroJerk-XYZ
tBodyAccMag
tGravityAccMag
tBodyAccJerkMag
tBodyGyroMag
tBodyGyroJerkMag
fBodyAcc-XYZ
fBodyAccJerk-XYZ
fBodyGyro-XYZ
fBodyAccMag
fBodyAccJerkMag
fBodyGyroMag
fBodyGyroJerkMag

The set of variables that were estimated from these signals are: 

mean(): Mean value
std(): Standard deviation
mad(): Median absolute deviation 
max(): Largest value in array
min(): Smallest value in array
sma(): Signal magnitude area
energy(): Energy measure. Sum of the squares divided by the number of values. 
iqr(): Interquartile range 
entropy(): Signal entropy
arCoeff(): Autorregresion coefficients with Burg order equal to 4
correlation(): correlation coefficient between two signals
maxInds(): index of the frequency component with largest magnitude
meanFreq(): Weighted average of the frequency components to obtain a mean frequency
skewness(): skewness of the frequency domain signal 
kurtosis(): kurtosis of the frequency domain signal 
bandsEnergy(): Energy of a frequency interval within the 64 bins of the FFT of each window.
angle(): Angle between to vectors.

Additional vectors obtained by averaging the signals in a signal window sample. These are used on the angle() variable:

gravityMean
tBodyAccMean
tBodyAccJerkMean
tBodyGyroMean
tBodyGyroJerkMean

The train and test set have 561 measurements.

## Getting data

The following code will download the data and unzip the file.

```{r}

## fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
## download.file(fileUrl,destfile="./acelerometers.zip")
## unzip("acelerometers.zip", exdir = ".")

```

## Obtaining the train set and test set

The following code will extract the train set and test set.

It will obtain other informations like the features list, the activitives list and labels.

```{r, warning=FALSE}

library(dplyr)

## Features list
features<-read.table("./UCI HAR Dataset/features.txt")
head(features)
dim(features)

## Activity labels
activity_labels<-read.table("./UCI HAR Dataset/activity_labels.txt")
print(activity_labels)

## Train set
train<-read.table("./UCI HAR Dataset/train/X_train.txt")
dim(train)
## Rename the variables for its descriptions
names(train)<-features[,2]

## Train set activities and subjects
ytrain<-read.table("./UCI HAR Dataset/train/y_train.txt")
table(ytrain)
## Changing the number of activity for its description
activity_train<-mutate(ytrain, V1 = activity_labels$V2[ytrain$V1])
table(activity_train)
## Rename the variable activity
names(activity_train)<-"activity"

strain<-read.table("./UCI HAR Dataset/train/subject_train.txt")
table(strain)
## ## Rename the variable subjec
names(strain)<-"subject"

## Merging train set, train activity and train subject
train<-cbind(train, activity_train, strain)
dim(train)

## Test set
test<-read.table("./UCI HAR Dataset/test/X_test.txt")
dim(test)
## Rename the variables for its descriptions
names(test)<-features[,2]

## Test set activities and subjects
ytest<-read.table("./UCI HAR Dataset/test/y_test.txt")
table(ytest)
## Changing the number of activity for its description
activity_test<-mutate(ytest, V1 = activity_labels$V2[ytest$V1])
table(activity_test)
## Rename the variable activity
names(activity_test)<-"activity"

stest<-read.table("./UCI HAR Dataset/test/subject_test.txt")
table(stest)
## ## Rename the variable subjec
names(stest)<-"subject"

## Merging test set, test activity and test subject
test<-cbind(test, activity_test, stest)
dim(test)

```


## Cleaning the data

In this section we will execute the activities asked.

### Merging train and test set

The train set and test set will be merged. Also, the variables will be renamed.

```{r}

dataset<-rbind(train,test)
dim(dataset)

```


### Extracting only the mean and standard deviation for each measurement

First we will find among all features the ones that evaluate mean and standard deviation of each measurement.

```{r}

## Mean subset among features
mean_subsets<-grep("mean\\()", features$V2)
print(mean_subsets)
head(features[mean_subsets,])

## Standar deviation subset among features
sd_subsets<-grep("std\\()", features$V2)
print(sd_subsets)
head(features[sd_subsets,])

```

Now, the complete dataset will be subseted and will remain with only mean and standard variation measurements.

```{r}

dataset_sub<-cbind(dataset[,mean_subsets], dataset[,sd_subsets], dataset[, 562:563])
dim(dataset_sub)

```


## New data set based in this one created

The final task is to create a new dataset with means of all measurements grouped by subject and activity.

```{r}

newdataset<-dataset_sub %>%
  group_by(activity, subject) %>%
    summarise_all(mean)

dim(newdataset)

## Write the final tidy data
write.table(newdataset, "tidydata.txt", row.name=FALSE)

```

